---
title: "Data from open data Toronto on Covid-19 and the amount of affected people in Toronto"
author: "Swarnadeep Chattopadhyay"
thanks: "https://github.com/WanteXecutioN/Sta304-Assignment-"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

```{r, include=FALSE}
library(opendatatoronto)
library(dplyr)
 
# get package
package <- show_package("64b54586-6180-4485-83eb-81e8fae3b8fe")
package
 
# get all resources for this package
resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
 
# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
covid_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
covid_data
```

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)

#removing the spaces in column headings
names(covid_data) <- str_replace_all(names(covid_data), c(" " = "." , "," = "" ))

#Top Toronto Neighbourhoods case count
nbr_ct <- data.frame(sort(table(covid_data$Neighbourhood.Name), decreasing = TRUE))
knitr::kable(head(nbr_ct, 10), "pipe", col.name =c("Neighbourhood Name", "Number of Cases"), align = c("l", "c"))

knitr::kable(head(covid_data, 5), "pipe")

#Create a copy of original data to add columns for Year and Year & Month
c19_cpy <- covid_data
c19_cpy$Yr <- format(as.Date(covid_data$Reported.Date), "%Y")
c19_cpy$Yr_Mth <- format(as.Date(covid_data$Reported.Date), "%Y-%m")

#Case Count by year
ggplot(c19_cpy, aes(Yr, ..count..))+
  geom_bar(aes(fill = Yr), alpha = 0.5)+
  geom_text(aes(label = after_stat(count)), vjust = 0, stat = "count")+
  labs(x = "Year", y = "Number of Cases", title = "YTD Numbers")

#Case Count by Months
Yr_20 <- c19_cpy %>% 
          filter(Yr == "2020") %>% 
          ggplot(aes(Yr_Mth, ..count..))+
          geom_bar(aes(fill = Yr_Mth), alpha = 0.5)+
          labs(x = "Months", y = "Number of Cases", title = "Monthly Numbers")+
          theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

Yr_21 <- c19_cpy %>% 
          filter(Yr == "2021") %>% 
          ggplot(aes(Yr_Mth, ..count..))+
          geom_bar(aes(fill = Yr_Mth), alpha = 0.5)+
          labs(x = "Months", y = "Number of Cases", title = "Monthly Numbers")+
          theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

Yr_22 <- c19_cpy %>% 
          filter(Yr == "2022") %>% 
          ggplot(aes(Yr_Mth, ..count..))+
          geom_bar(aes(fill = Yr_Mth), alpha = 0.5)+
          labs(x = "Months", y = "Number of Cases", title = "Monthly Numbers")+
          theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

ggarrange(Yr_20, Yr_21, Yr_22,
          labels = c("2020", "2021", "2022"),
          nrow = 3)

#Total case count by age groups
ggplot(covid_data, aes(Age.Group, ..count..))+
  geom_bar(aes(fill = Age.Group), alpha = 0.5)+
  labs(x = "Age Groups", y = "Number of Cases", title = "Cases by Age groups")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

#leading factors for covid infections
ggplot(covid_data, aes(Source.of.Infection, ..count..))+
  geom_bar(aes(fill = Source.of.Infection), alpha = 0.5)+
  labs(x = "Sources", y = "Number of Cases", title = "Factors leading to cases")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+
  coord_flip()
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 



# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


